Towards a Classification of Isolated $j$-invariants

### 2023-05-18 TeX/draft.tex
Updates from Overleaf

\documentclass[11pt,reqno]{amsart}
% We ain't got no time for eq. nums. on the left

@@

% Presented to you by Technicolor, and the number 3
\usepackage{graphics}
\usepackage{hyperref}
\usepackage[usenames, dvipsnames]{xcolor}

@@

% Nongross real and imaginary parts
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\Q}{\bQ}
\newcommand{\Z}{\bZ}
\newcommand{\PP}{\mathbb P}
\newcommand{\filip}[1]{{\color{cyan} \textsf{$\sun\sun\sun$ Filip: [#1]}}}
\newcommand{\davidnote}[1]{\textcolor{Plum}{David: #1}}

\newcommand{\abbey}[1]{\textcolor{blue}{Abbey: #1}}
\newcommand{\dld}[1]{\textcolor{Plum}{dld: #1}}
\newcommand{\sachi}[1]{\textcolor{purple}{Sachi: #1}}
\newcommand{\timo}[1]{\textcolor{red}{Timo: #1}}
% \newcommand{\travis}[1]{\textcolor{chartreuse}{Travis: #1}}

% Don't have subsections appear in TOC
%\setcounter{tocdepth}{1}

\title{Towards a Classification of Sporadic $j$-invariants}
\input{author_info.tex}
\date{\today}

\begin{document}
\begin{abstract}
We develop an algorithm to test whether a non-CM elliptic curve $E/\Q$ gives rise to a sporadic point of any degree on any modular curve of the form $X_1(N)$. This builds on prior work of Zywina which gives a method for computing the image of the adelic Galois representation associated to $E$. Running this algorithm on all elliptic curves presetly in the LMFDB yields strong evidence for the fact that $E$ gives rise to a sporadic point if and only if $j(E)=-140625/8$ or $-9317$.
    \end{abstract}
\maketitle

The modular curve $X_1(N)$ is an algebraic curve over $\Q$ whose non-cuspidal points parametrize elliptic curves with a distinguished point of order $N$. We are interested in studying \textbf{sporadic} points $x \in X_1(N)$, which are points for which there are only finitely many points on $X_1(N)$ of degree at most $\deg(x)$. Hence a non-cuspidal sporadic point corresponds to an elliptic curve with a point of order $N$ in ``usually low degree." Elliptic curves with complex multiplication (CM) provide many natural examples of sporadic points, as the extra endomorphisms of a CM elliptic curve give constraints on the size of the image of the associated Galois representation. Indeed, there exist sporadic CM points on $X_1(N)$ for all $N\geq 721$; see \cite[Theorem 8.2]{CGPS2022}.

@@

\section{Background}

\begin{theorem}
Let $E/\Q$ be a non-CM elliptic curve, and let $\rho_E$ denote the adelic Galois representation of $E$. Let $M_E$ be any positive integer such that
\[
\im \rho_E=\pi^{-1}(\im \rho_{E, M_E}).
\]
Then for any $x \in X_1(n)$ with $j(x)=j(E)$ we have $\deg(x)=\deg(f)\cdot \deg(f(x))$, where $f: X_1(n) \rightarrow X_1(\gcd(n,M_E))$ denotes the natural map. It follows that:
\begin{enumerate}
\item If $x$ is $\mathbf{P}^1$-isolated, then $f(x)$ is $\mathbf{P}^1$-isolated.
\item If $x$ is AV-isolated, then $f(x)$ is AV-isolated.
\item If $x$ is sporadic, then $f(x)$ is sporadic.
\end{enumerate}
\end{theorem}

@@

\section{Proving that a point is not sporadic}
The following algorithm is the main procedure for determining a $j$-invariant is not sporadic. It gives an overview of the structure.

\begin{algorithm}[h!]
\caption{Main Algorithm}\label{alg:mainalgorithm}
\KwIn{A non-CM $j$-invariant $j = j(E) \in \Q$ and $a$-invariants for $E$ with $j$-invariant $j$.}
\KwOut{Finite list of pairs $(n,d)$ such that any isolated point on $X_1(N)$ associated to $E$ maps down under the natural projection map to an isolated point of degree $d$ on $X_1(n)$ in the list. }
Compute the adelic image $G$ of $E/\Q$ as a subgroup of $\GL_2(\hat{\Z})$ using Zywina's algorithm. Represent the output as some \emph{level} $N$ together with a subgroup $G$ of $\GL_2(\Z/N\Z)$\;
Using Algorithm \ref{alg:reduce_level}, reduce the level to $G, m_0$\; 
Define $L \colonequals \{\}$\;
\For{all $12 < b \mid m_0$}{
Using Algorithm~\ref{alg:compute_degree} compute the set $D$ of degrees of all fields of definition of a point with respect to $G \mod b$, the image of $G$ in $\GL_2(\Z/b\Z)$\;
Set $L \colonequals L \cup \{ (b, d) : d \in D\}$ \;}
Using Algorithm \ref{alg:level_mapping} filter the list $L$ to obtain a refined list $L'$\;
\Return{the list $L'$ } \;
    %\filip{We need to look at the orbits of $\langle G, -G \rangle \mod b$ on $(\Z/b\Z)^2$, not $\bP^1(\Z/b\Z)$. If we consider the action on $\bP^1(\Z/b\Z)$, the orbits might be too small and hence the degree we get. Furthermore, it is necessary to consider only the elements of $(\Z/b\Z)^2$ of order $b$.}
\end{algorithm}

\begin{theorem}
 Algorithm \ref{alg:mainalgorithm} is correct. 
\end{theorem}
\begin{corollary}
If Algorithm \ref{alg:mainalgorithm} outputs an empty set on some $j$-invariant $j = j(E)$ and $a$-invariants $a$ for $E$, then $j$ is not a  $\mathbf{P}^1$-isolated point on $X_1(N)$.
\end{corollary}
\abbey{Zywina said the results returned from his adelic image algrithm are guaranteed to be correct and that ``errors will always occur if $E$ gives rise to an unknown exceptional rational point on certain high genus modular curves." Since an error in his algorithm will result in an error in our algorithm, I don't know what should be said about this in our writeup.} \sachi{maybe the easiest thing is just to add a remark saying that the results are guaranteed to be correct, and getting an error instead of a result is possible but that we did not encounter it?} \timo{I second that.} \abbey{alternatively, separate Zywina's algorithm from ours in this paper}

@@

\section{Reduce Level}

\abbey{Travis,Sachi, Abbey}

Using Zywina's algorithm for the level of the adelic Galois representation as a starting point, we may obtain the level of the $m$-adic Galois representation associated to $E$ as in Theorem \ref{BELOVthm}.

\begin{algorithm}[H]
  \KwIn{$G\subseteq \GL_2(\Z/N\Z)$ such that $\im\rho_E=\pi^{-1}(G)$ and $N$ is the level}
  \KwOut{$G\subseteq \GL_2(\Z/m_0\Z)$ such that $G=\im\rho_{E,m_0}$ and $\im\rho_{E,m^{\infty}}=\pi^{-1}(\im\rho_{E,m_0})$ where $m$ is the product of the non-surjective primes and $2$ and $3$ (if they do not already divide $m$), and $m_0$ is minimal }
  \caption{Reduce level}%
  \label{alg:reduce_level}
  Let $P_E = \{\ell|N \text{ prime} \colon \rho_{E,\ell} \text{ is non-surjective} \}$\;
  Let $S_E= P_E \cup \{ 2, 3\}$\;
  Let $m_0 = \prod_{\ell \in S_E} \ell^{v_{\ell}(N)}$\;
  \For{$p \in \{ \ell | m_0 \colon \ell \text{ prime} \}$}{
  \While{$p |m_0$ and $\#G/\#(G \pmod{m_0/p}) = p^4$}{
    $m_0 = m_0/p$\;
    $G = G \pmod{m_0}$\;
  }
  \If{not $p$ in $P_E$ and $\ord_p(m_0) = 1$ and $\#G/\#(G \pmod{m_0/p}) = \#\GL_2(\Z/p\Z)$}{
  $m_0 = m_0/p$\;
   $G = G \pmod{m_0}$\;}}
  \Return{$G,m_0$}
\end{algorithm}

  \begin{proposition}
    Algorithm~\ref{alg:reduce_level} is correct.
  \end{proposition}
  \begin{proof}
    Since $N$ is the level of the adelic representation and both $m$ and $m_0$ divides $N$, we must have that $\im\rho_{E, m^\infty} = \pi^{-1}(\im \rho_{E,m_0})$. This property is an invariant of the for-loop. The for-loop
    terminates once as many powers of $\ell$ have been removed from $m_0$
    as possible while maintaining this invariant. Thus $m_0$ is the minimal such integer and is thus the level.
  \end{proof}

@@

The following algorithm computes the degrees of closed points on $X_1(N)$ associated to $E$.

>>>>

\documentclass[11pt,reqno]{amsart}
% We ain't got no time for eq. nums. on the left
\usepackage{soul}%added for \st

@@

% Presented to you by Technicolor, and the number 3
\usepackage{graphics}
%\usepackage{ulem}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage[usenames, dvipsnames]{xcolor}

@@

% Nongross real and imaginary parts
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\Q}{\bQ}
\newcommand{\C}{\bC}
\newcommand{\Z}{\bZ}
\newcommand{\PP}{\mathbf P}
\newcommand{\Jac}{\operatorname{Jac}}
\newcommand{\proj}{\operatorname{proj}}

\newcommand{\davidnote}[1]{\textcolor{Plum}{David: #1}}
\newcommand{\filip}[1]{{\textcolor{cyan}{Filip: [#1]}}}
\newcommand{\abbey}[1]{\textcolor{blue}{Abbey: #1}}
\newcommand{\dld}[1]{\textcolor{Plum}{dld: #1}}
\newcommand{\sachi}[1]{\textcolor{purple}{Sachi: #1}}
\newcommand{\timo}[1]{\textcolor{red}{Timo: #1}}
\newcommand{\travis}[1]{\textcolor{orange}{Travis: #1}}
\newcommand{\abedit}[1]{{\color{blue} #1}}
\newcommand{\fedit}[1]{{\color{cyan} #1}}
% Don't have subsections appear in TOC
%\setcounter{tocdepth}{1}

\title{Towards a Classification of Isolated $j$-invariants}
\input{author_info.tex}
\date{\today}

\begin{document}
\begin{abstract} \abbey{will update once we re-run the code}
We develop an algorithm to test whether a non-CM elliptic curve $E/\Q$ gives rise to a sporadic point of any degree on any modular curve of the form $X_1(N)$. This builds on prior work of Zywina which gives a method for computing the image of the adelic Galois representation associated to $E$. Running this algorithm on all elliptic curves presetly in the LMFDB yields strong evidence for the conjecture that $E$ gives rise to a sporadic point if and only if $j(E)=-140625/8$ or $-9317$.
    \end{abstract}
\maketitle
\abbey{will update once we re-run the code}

The modular curve $X_1(N)$ is an algebraic curve over $\Q$ whose non-cuspidal points parametrize elliptic curves with a distinguished point of order $N$. We are interested in studying \textbf{sporadic} points $x \in X_1(N)$, which are points for which there are only finitely many points on $X_1(N)$ of degree at most $\deg(x)$. Hence a non-cuspidal sporadic point corresponds to an elliptic curve with a point of order $N$ in ``unusually low degree." Elliptic curves with complex multiplication (CM) provide many natural examples of sporadic points, as the extra endomorphisms of a CM elliptic curve give constraints on the size of the image of the associated Galois representation. Indeed, there exist sporadic CM points on $X_1(N)$ for all $N\geq 721$; see \cite[Theorem 8.2]{CGPS2022}.

@@

\section*{Acknowledgments}
We thank Pete Clark, Jeremy Rouse, and David Zureick-Brown for helpful conversations.

\section{Background}
\subsection{Isolated Points on Curves}
Let $C$ be a curve, by which we will mean a nonsingular projective geometrically integral 1-dimensional scheme defined over a number field $k$; we suppose all curves satisfy these assumptions throughout the paper. To streamline our exposition, we assume $P_0 \in C(k)$, but this is not required; see \cite[$\S4$]{BELOV}. Throughout, we consider closed points of the curve $C$, which correspond to $\Gal_k$-orbits of points in $C(\overline{k})$. The degree of $x$ is defined to be the degree of the residue field $k(x)$ over $k$, or alternatively, to be the length of the $\Gal_k$-orbit of points in $C(\overline{k})$ corresponding to $x$.

To any closed point $x \in C$ of degree $d$ we associate the $k$-rational effective divisor 
\[
 x=P_1+ \cdots +P_{d},
\]
where $P_1, \dots, P_{d}$ are the points in the $\Gal_{k}$-orbit associated to $x$. With this identification, we can study the image of $x$ under the natural map from the $d$th symmetric power of $C$ to the curve's Jacobian
\[
\Phi_d: C^{(d)} \rightarrow \Jac(C)
\]
which sends the effective divisor $D$ of degree $d$ to the class $[D - dP_0]$, where $P_0 \in C(k)$. If $\Phi_d(x)=\Phi_d(y)$ for some other point $y \in C^{(d)}(k)$, then there exists a non-constant function $f \in k(C)^{\times}$ with $\text{div}(f)=x-y$. Since $x$ is a degree $d$ point and $x\not=y$, $x$ and $y$ have distinct support so $f: C \rightarrow \mathbb{P}^1_{k}$ is a dominant morphism of degree $d$.\footnote{In particular, this shows $\Phi_d$ is injective if $d$ is less than the $k$-gonality of $C$, which is the least degree of a non-constant rational map to $\mathbb{P}^1$.} By Hilbert's irreducibility theorem \cite[Chapter 9]{serre97}, $f^{-1}(\mathbb{P}^{1}(k))$ will contain infinitely many closed points of degree $d$. On the other hand, if $\Phi_d$ is injective on closed points of degree $d$, then by Faltings' Theorem \cite{faltings} there can exist infinitely many such points only if all but finitely many are parametrized by translates of positive rank abelian subvarieties of $\Jac(C)$. This inspires the following:

\begin{definition} Let $C$ be a curve defined over a number field $k$.
\begin{enumerate}
\item []
\item A closed point $x \in C$ of degree $d$ is $\PP^1$\textbf{-parametrized} if there exists a point $x' \in C^{(d)}(k)$ such that $\Phi_d(x)=\Phi_d(x')$.
\item A closed point $x \in C$ of degree $d$ is \textbf{AV-parametrized} if there exists a positive rank abelian subvariety $A/k$ with  $A \subset \Jac(C)$ such that $\Phi_d(x)+A \subset \im(\Phi_d)$.
\item A closed point $x \in C$ of degree $d$ is \textbf{isolated} if it is neither $\mathbb{P}^1$-parametrized nor AV-parametrized.
\item A closed point $x \in C$ of degree $d$ is \textbf{sporadic} if there are only finitely many closed points of $C$ of degree at most $\deg(x)$.
\end{enumerate}
\end{definition}

If $C$ has genus $g\geq 2$, then the collection of all points on $C$ with coordinates in $k$ is finite by Faltings's theorem \cite{faltings83}. In general, the set $C(k)$ sits inside a larger finite set of points, namely, the set of all isolated points of $C$.
    \begin{theorem}[{Bourdon, Ejder, Liu, Odumodu, Viray, \cite[Theorem 4.2]{BELOV}}]\label{thm:FiniteIsolated}
        Let $C$ be a curve over a number field.
        \begin{enumerate}
            \item There are infinitely many degree $d$ points on $C$ if and only if there is a degree $d$ point on $C$ that is \emph{not} isolated.
            \item There are only finitely many isolated points on $C$.
        \end{enumerate}
    \end{theorem}
It follows from Theorem \ref{thm:FiniteIsolated} that every sporadic point is isolated, but the converse need not hold.

\subsection{Modular Curves}
For any subgroup $H \leq \GL_2(\Z/N\Z)$, we define the modular curve $X_H$ to be the coarse space of the stack $\mathcal{M}_H$, as defined in Deligne--Rapoport \cite{DR}. \abbey{Describe interpretation of $K$-rational points.} The curve $X_H$ is a scheme over $\text{Spec }\Z[1/N]$ and parametrises generalized elliptic curves with $H$-level structure. See, for example, \cite[$\S2.3$]{RSZB2022} for details. In particular, if we take
\[
B_1(N) =\left\{ \begin{pmatrix}
1 & * \\
0 & *
\end{pmatrix} \right \} \leq \GL_2(\Z/N\Z),
\]
then $X_{B_1(N)}=X_1(N)$, the modular curve whose noncuspidal points parametrize  elliptic curves with a distinguished point of order $N$. There is an analytic isomorphism between $X_1(N)(\mathbb{C})$ and the Riemann surface constructed as a quotient of the extended upper-half plane by the congruence subgroup
\[
\Gamma_1(N) \coloneqq\left\{ \begin{pmatrix}
a & b \\
c & d 
\end{pmatrix} \in \SL_2(\Z) \, : \, c \equiv 0 \pmod{N} \text{ and } a \equiv d \equiv 1 \pmod{N} \right \}, 
\]
with matrices acting via linear fractional transformations. If $N \geq 4$, then $\mathcal{M}_{B_1(N)}$ is its own coarse moduli space, and so noncuspidal $k$-rational points of $X_1(N)$ classify pairs $(E,P)$, where $E/k$ is an elliptic curve and $P\in E(k)$, up to $k$-isomorphism. \abbey{I decided not to define the modular curve by its moduli problem as in \cite[$\S2$]{RouseDZB}, since the construction yields a stack in general... so it seems confusing to call it a modular *curve*. I've opted to go with the conventions of \cite[$\S2.3$]{RSZB2022}.}

\abbey{maybe define $X_G$ for general open subgroup $G \leq \GL_2(\widehat{\Z})$ here instead of in later section on maps between modular curves}

\subsection{Closed Points on Modular Curves} To discuss isolated points on modular curves, we must consider closed points on $X_1(N)$, viewed always as a scheme over $\Q$. \fedit{Let $k$ be a number field and fix an embedding of $k$ into $\overline \Q$}.
Given an elliptic curve $E/k$ with $P \in E(k)$ of order $N$, the pair $(E,P)$ induces a $k$-valued point on $X_1(N)$ via the moduli interpretation described above. We denote this $k$-valued point by $(E,P)_k$. %\abedit{and it corresponds to a morphism of $\Q$-schemes from $\text{Spec } k$ to $X_1(N)$. 
\fedit{The $\Gal_\Q$-orbit of $(E,P)_k$ corresponds bijectively to a closed point on $X_1(N)$ over $\Q$, and we call this the \textbf{closed point} $x \in X_1(N)$} associated to $(E,P)$ \abbey{I much prefer the earlier definition -- the bijection of closed points with Galois orbits is something that requires proof, whereas what I wrote originally comes directly from the definition of $k$-valued point, i.e, a morphism from $\text{Spec }k$ to $X$. This makes me disagree with your later edits to this section.} \filip{I think the bijection is obvious and can in any case easily be cited, e.g. from \cite[Proposition 2.4.6]{Poonen_RPV} and makes everything clearer. But if you insist, feel free to revert to what you previously wrote},  and we define the \textbf{degree} of $x$ to be the degree of the residue field $\Q(x)$ over $\Q$.

We note in particular that $\deg(x)$ may be strictly less than the degree of $k$. However, there always exists $E'/\Q(x)$ with $j(E')=j(E)$ and $P' \in E'(\Q(x))$, where the point $P' \in E'$ maps to $P \in E$ under a $\overline \Q$-isomoprhism sending $E$ to $E'$. See \cite[p. 274, Proposition VI.3.2]{DR}. \fedit{The pair $(E',P')$ gives a $\Q(x)$-valued point \st{associated to the closed point $x$}, and it is the unique $\Q(x)$-valued point such that $(E,P)_{k}=(E',P')_{k}.$}

In applications where such a property is important, it is sometimes useful to work with the \textbf{geometric point} associated to $(E,P)$, which we define to be the point $(E,P)_{\overline{\Q}}$. Note $(E,P)_{\overline{\Q}}$ induces the same closed point as $(E,P)_k$.

Let $E/k$ be an elliptic curve and $P \in E(k)$ a point of order $N$. Note that for any $\xi \in \Aut(E)$, the pair $(E,\xi P)$ induces the same closed point $x \in X_1(N)$, since $\xi$ provides the necessary isomorphism.\footnote{ \abbey{Add footnote about connection with $X_{\pm B_1(N)}$. ``If you take the moduli space of the set of closed points associated to pairs $(E,P)$ as above, you in fact get exactly $X_{\pm B_1(N)}$ (i.e the moduli functor will correspond exactly to this)!" Does this still hold for $j=0,1728$?}} This can be used to obtain a more explicit description of the residue field $\Q(x)$.

\begin{lemma}\label{ResidueFieldLemma}
Let $E$ be an elliptic curve defined over $\Q(j(E))$, and let $P \in E$ be a point of order $N$. Then the residue field of the closed point $x \in X_1(N)$ associated to $(E,P)$ is given by
\[
\Q(x)=\Q(j(E),\mathfrak{h}(P)),
\]
where $\mathfrak{h} \rightarrow E/\Aut(E) \cong \mathbb{P}^1$ is a Weber function for $E$. \abbey{Need to clarify statement to match conventions about $k$-valued pts/closed points}
\end{lemma}

\begin{proof}
See, for example, \cite[Lemma 2.5]{OddDegQCurve}. 
\end{proof}

\begin{remark} In particular, if $E/\Q(j(E))$ does not have complex multiplication and $P=(x_0,y_0)$, then one can take $\Q(x)=\Q(j(E),x_0)$. See \cite[p. 107]{shimura}.
\end{remark}
\begin{example}
Let $E_1:y^2+xy+y=x^3-x^2-3x+3$ and $P_1=(-1,2)$ a point of order 7. Then $(E_1,P_1)$ gives a $\Q$-valued point on $X_1(7)/\Q$ and also a closed point $x\in X_1(7)$ of degree 1. On the other hand, let $E_2:y^2=x^3-43x-166$ and $P_2=(5,i\sqrt{256})$ a point of order 7. Then $(E_2,P_2)$ gives a $\Q(i)$-valued point on $X_1(7)/\Q$ and also a closed point $x\in X_1(7)$ of degree 1 by Lemma \ref{ResidueFieldLemma} \abedit{since $\Aut(E) = \{\pm1\}$}. In fact, both $(E_1,P_1)$ and $(E_2,P_2)$ induce the same geometric point on $X_1(7)$ since $(E_1,P_1)_{\overline{\Q}}=(E_2,P_2)_{\overline{\Q}}$. One could alternatively compute the Kubert-Tate normal form associated to $E_2/\Q(i)$, with $P_2=(5,i\sqrt{256})$:
\[E_3: y^2-xy-4y=x^3-4x^2,\, \, P_3=(0,0).
\]
We can check that $(E_1,P_1)_{\Q}=(E_3,P_3)_{\Q}$ and $(E_2,P_2)_{\Q(i)} =(E_3,P_3)_{\Q(i)}$. Thus it is fair to say that $(E_2,P_2)$ induces a $\Q$-valued point, even though it is not itself a $\Q$-valued point. 
\end{example}

\begin{example}
The distinction between closed points and geometric points can be seen when counting the number of points of a particular degree. For example, let $E$ be the elliptic curve with LMFDB label \href{https://www.lmfdb.org/EllipticCurve/Q/162/c/3}{162.c3}. Then $E$ possesses a unique $\Q$-rational subgroup of order 21 with generator $P$. For each $a \in (\Z/21\Z)^{\times}$, we consider the geometric point on $X_1(21)$ associated to $(E,aP)$. Since $(E,aP)_{\overline{\Q}}=(E,-aP)_{\overline{\Q}}$, we find there are six distinct geometric points corresponding to $(E,aP)$ for $a \in (\Z/21\Z)^{\times}$. However, these six $\overline{\Q}$-points lie in two Galois orbits, each of size 3. Thus we would say there are two closed points of degree 3, and the cardinality of the Galois orbit equals the degree.
\end{example}

%\begin{example}
%\emph{To add; could also talk about $f^*(j(E))$}
%\end{example}

\subsection{Maps Between Modular Curves} We say an open subgroup $G$ of $\GL_2(\widehat{\Z})$ has \textbf{level} $N$ if $G=\pi^{-1}(\overline{G})$, where $\pi:\GL_2(\widehat{\Z}) \rightarrow \GL_2(\Z/N\Z)$ is the natural projection map and $\overline{G}=\pi(G)$. \timo{and $N$ is minimal with respect to this property} If $\det(G)=\widehat{\Z}^{\times}$, we define the modular curve $X_G \coloneqq X_{\overline{G}}$. If $G=\GL_2(\widehat{\Z})$, then we identify $X_G \cong \mathbb{P}^1$ with the $j$-line.
\begin{proposition} \label{prop:mapdegree}
If $G \subseteq G' \subseteq \GL_2(\widehat{\Z})$ are two open subgroups with surjective determinant, then there is a natural $\Q$-rational morphism $X_G \rightarrow X_{G'}$ of degree $[\pm G': \pm G]$. \timo{where $\pm G$ is the subgroup generated by $G$ and $-I_2$}
\end{proposition}
\begin{proof} This comes from the inclusion of function fields $\Q(X_G) \supseteq \Q(X_{G'})$.
\abbey{ Is there a good citation for this? Zywina's paper just states it, and other papers give a more precise justification, but work with subgroups of $\GL_2(\Z/N\Z)/\{\pm 1\}$. We should also mention why surjective determinant is needed here... my understanding is that if the determinant is not surjective, then the function field construction will not result in a curve over $\Q$, but what we define to be $X_H$ is always a scheme over $\Q$. So the two constructions only coincide exactly when the determinant is surjective. I think it's worth mentioning this somewhere, since Zywina's paper uses the function field construction.} \timo{My guess is that it would follow from the construction of $X_G$ as a quotient, probably with additional work for going from the fine to the coarse moduli space.}\\
\filip{Here's a  proof: Let $N$ be the level of $G$. For any subgroup $H$ of $\GL_2(\Z /N\Z)$, we have $\Q(X_H)=\Q(X(N))^H$. Now by Galois theory it follows that $\Q(X_G) \supseteq \Q(X_{G'})$, so we conclude that there exists a $\Q$-rational morphism $f:X_G \rightarrow X_{G'}$.\\ To determine its degree, let $\Gamma$ and $\Gamma'$ be the intersection with $\SL_2(\Z)$ of the inverse image of $G$ and $G'$ in $\GL_2(\Z)$. Over $\bC$, the morphism $f$ is the quotient map $\Gamma \backslash \mathcal{H} \rightarrow \Gamma' \backslash \mathcal{H}$. Since the kernel of the action of $\SL_2(\Z)$ on $\mathcal H$ is $\pm I$, the degree of $f$ is as claimed.}
    \end{proof}
Taking $G=\pi^{1}(B_1(ab))$ and $G'=\pi^{-1}(B_1(a))$ gives the following corollary, which under the moduli interpretation corresponds to sending $(E,P)$ to $(E,bP)$.    \begin{corollary}
                For positive integers $a$ and $b$, the natural $\Q$-rational map $f\colon X_1(ab) \rightarrow X_1(a)$ has
                \[
                    \deg(f)=
                    c_{f}\cdot b^2 \prod_{p \mid b,\, p \nmid a}
                    \left(1-\frac{1}{p^2}\right),
                \]
                where $c_{f}=1/2$ if $a \leq 2$ and $ab>2$ and $c_{f}=1$ otherwise. 
    \end{corollary}

\subsection{Galois Representations} \abbey{still need to add}
\[
\rho_{E,m^{\infty}}: \Gal_k \xrightarrow{\rho_E} \GL_2(\widehat{\Z}) \cong  \prod_{p \text{ prime}} \GL_2(\Z_{p}) \xrightarrow{\proj} \prod_{p \mid m} \GL_2(\Z_{p}),
\]

\section{Isolated Points on Modular Curves: Prior Results}
\abbey{Not sure which results need to be included, or if they should be discussed in the introduction.}\\
\abbey{Need to define isolated/sporadic $j$-invariant; maybe in intro}
\begin{theorem}
Let $E/\Q$ be a non-CM elliptic curve, and let $\rho_E$ denote the adelic Galois representation of $E$. Let $M_E$ be any positive integer such that
\[
\im \rho_E=\pi^{-1}(\im \rho_{E, M_E}).
\]
Then for any $x \in X_1(n)$ with $j(x)=j(E)$ we have $\deg(x)=\deg(f)\cdot \deg(f(x))$, where $f: X_1(n) \rightarrow X_1(\gcd(n,M_E))$ denotes the natural map. \timo{Separate this into a corollary:?} It follows that:
\begin{enumerate}
\item If $x$ is $\mathbf{P}^1$-isolated, then $f(x)$ is $\mathbf{P}^1$-isolated.
\item If $x$ is AV-isolated, then $f(x)$ is AV-isolated.
\item If $x$ is sporadic, then $f(x)$ is sporadic.
\end{enumerate}
\end{theorem}

@@

\section{Proving that a $j$-invariant is not isolated}
The following algorithm is the main procedure for determining a $j$-invariant is not isolated. It gives an overview of the structure.

\begin{algorithm}[h!]
\caption{Main Algorithm}\label{alg:mainalgorithm}
\KwIn{A non-CM $j$-invariant $j = j(E) \in \Q$.}
\KwOut{A finite list of pairs $(n,d)$ such that any isolated point on $X_1(N)$ associated to $E$ maps down under the natural projection map to an isolated point of degree $d$ on $X_1(n)$ in the list.}
Compute the adelic image $G$ of $E/\Q$ as a subgroup of $\GL_2(\hat{\Z})$ using Zywina's algorithm. Represent the output as some \emph{level} $N$ together with a subgroup $G$ of $\GL_2(\Z/N\Z)$\;
Using Algorithm \ref{alg:reduce_level}, reduce the level to $G, m_0$\; 
Set $L \colonequals \{\}$\;
\For{all $12 < b \mid m_0$}{
Using Algorithm~\ref{alg:compute_degree} compute the set $D$ of degrees of all fields of definition of a point with respect to $G \mod b$, the image of $G$ in $\GL_2(\Z/b\Z)$\;
Set $L \colonequals L \cup \{ (b, d) : d \in D\}$\;}
Using Algorithm \ref{alg:level_mapping} filter the list $L$ to obtain a refined \timo{reduced?} list $L'$\;
\Return{the list $L'$ }\; \abbey{Appears to also return true/false. I think the false output is misleading since false does not mean ``not isolated," only that we cannot rule it out} \sachi{Perhaps this is a case where it's clearer if our code and our paper do not match completely. For me the true/false was so that I can easily parse the data and find out which ones need extra work. False just meant ``not done with this yet''. But I can see an argument also that it's misleading to the user to return a boolean on a function called notisolated... where the boolean does not indicate whether the point is isolated/not.}
\end{algorithm}

\begin{theorem}
 Algorithm \ref{alg:mainalgorithm} is correct. 
\end{theorem}
\abbey{proof needed somewhere -- should it go at the end after the other parts of the algorithm have been described?}
\begin{corollary}
If Algorithm \ref{alg:mainalgorithm} outputs an empty set on some $j$-invariant $j = j(E)$, then $j$ is not a  $\mathbf{P}^1$-isolated point on $X_1(N)$. \abbey{Actually, can conclude not isolated $j$-invariant}
\end{corollary}
\abbey{Zywina said the results returned from his adelic image algrithm are guaranteed to be correct and that ``errors will always occur if $E$ gives rise to an unknown exceptional rational point on certain high genus modular curves." Since an error in his algorithm will result in an error in our algorithm, I don't know what should be said about this in our writeup.} \sachi{maybe the easiest thing is just to add a remark saying that the results are guaranteed to be correct, and getting an error instead of a result is possible but that we did not encounter it?} \timo{I second that.} \abbey{alternatively, separate Zywina's algorithm from ours in this paper}

@@

\section{Reduce Level}
\abbey{I updated this section to match the revised code and implement other expository changes.}
Let $E/\Q$ be a non-CM elliptic curve, and define the set 
\[
S_E\coloneqq \{2,3\} \cup \{ \ell: \rho_{E,\ell^{\infty}}(\Gal_{\Q}) \neq \GL_2(\Z_{\ell}) \}.
\] For $m \coloneqq \prod_{\ell \in S_E} \ell$, we give an algorithm to obtain the level of the $m$-adic Galois representation associated to $E$ from the image of the adelic representation of $E$. In practice, we can obtain $\im \rho_E$ from Zywina's algorithm \cite{ZywinaAlgorithm}.

\abbey{Rewrite algorithm description to be less technical? Could say the algorithm finds the smallest $m \mid N$ such that $\#G=\#G_m \cdot \#\ker(\GL_2(\Z/mn\Z) \rightarrow \GL_2(\Z/m\Z))$ and returns $m=m0$.}

\begin{algorithm}[H]
  \KwIn{$G\leq \GL_2(\Z/N\Z)$ such that $\im\rho_E=\pi^{-1}(G)$ and $N$ is the level} 
  \KwOut{Smallest $m_0 \in \Z^+$ such that $\im\rho_{E,m^{\infty}}=\pi^{-1}(\im\rho_{E,m_0})$ where $m= \prod_{\ell \in S_E} \ell$}
  \caption{Reduce level}%
  \label{alg:reduce_level}
  Let $P_E = \{\ell \mid N \text{ prime} \colon \rho_{E,\ell} \text{ is non-surjective} \}$\;
  Let $S_E= P_E \cup \{ 2, 3\}$\;
  Let $m_0 = \prod_{\ell \in S_E} \ell^{v_{\ell}(N)}$\;
  \For{$p \in \{ \ell \mid  m_0 \colon \ell \text{ prime} \}$}{
  \While{$p \mid m_0$ and $\#G/\#(G \pmod{m_0/p}) = p^4$}{
    $m_0 = m_0/p$\;
    $G = G \pmod{m_0}$\;
  }
  \If{not $p$ in $P_E$ and $\ord_p(m_0) = 1$}{\If{$m_0=p$ and $\#G=\#\GL_2(\Z/p\Z)$}{\Return{1}}\ElseIf{$\#G/\#(G \pmod{m_0/p}) = \#\GL_2(\Z/p\Z)$}{
    $m_0 = m_0/p$\;
  } }
}
  %   \If{not $p$ in $P_E$ and $\ord_p(m_0) = 1$ and $\#G/\#(G \pmod{m_0/p}) = \#\GL_2(\Z/p\Z)$}{$m_0 = m_0/p$\;$G = G \pmod{m_0}$\;}
  \Return{$m_0$}
\end{algorithm}

  \begin{proposition}
    Algorithm~\ref{alg:reduce_level} is correct.
  \end{proposition}

  \begin{proof}
    Since $N$ is the level of the adelic representation and both $m$ and $m_0$ divides $N$ \abbey{It is not true that $m$ must $N$ since $m$ is always divisible by 2 and 3. But if $2$ does not divide $N$, for example, then it does not divide the level of the $m$-adic representation.}, we must have that $\im\rho_{E, m^\infty} = \pi^{-1}(\im \rho_{E,m_0})$. This property is an invariant of the for-loop. The for-loop
    terminates once as many powers of $\ell$ have been removed from $m_0$
    as possible while maintaining this invariant. Thus $m_0$ is the minimal such integer and is thus the level.
  \end{proof}

      \abbey{Perhaps we should have a proposition statement and proof that is independent of the algorithm writeup, and refer to this in the proof of the algorithm. A statement along the lines of: ``Let $G \leq \GL_2(\Z/mn\Z)$ and $G_m \coloneqq G \pmod{m}$. Suppose 
  \[
\#G=\#G_m \cdot \#\ker(\GL_2(\Z/mn\Z) \rightarrow \GL_2(\Z/m\Z)).
  \] Then $G= \pi^{-1}(G_m),$ where $\pi \rightarrow \GL_2(\Z/mn\Z) \rightarrow \GL_2(\Z/m\Z)$." But we'll need to address what's going on with 2 and 3 in particular.}

@@

The following algorithm computes the degrees of closed points on $X_1(N)$ associated to a \abedit{non-CM elliptic curve $E/\Q$}.
